---
title: "Covid in Victoria during 1st May to 30th September 2021"
subtitle: "Team omicron"
author: "Davide Signor, Ibrahim Al-Hindi, Janice Hsin Hsu"
institute: "Monash University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  
  chunk_output_type: console
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
        echo = FALSE,
        warning = FALSE, 
        message = FALSE,
        fig.retina = 5)
```


```{r loadlibrary, include=FALSE, echo=FALSE}
library(tidyverse)
library(xaringanthemer)
library(lubridate)
library(kableExtra)
library(sf)
library(plotly)
library(gganimate)
library(glue)
library(viridis)
library(ggthemes)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
style_mono_accent(
  base_color = "#fb8072",
  header_h1_font_size = "2.25rem",
  header_h2_font_size = "1.75rem",
  header_h3_font_size = "1.5rem",
  header_font_google = google_font("Pangolin"),
  text_font_google   = google_font("Pangolin"),
  text_font_size = "1.5rem",
  code_font_google   = google_font("Shadows Into Light Two"),
  code_font_size = "0.5rem"
) 

mypal <- c("#b2182b", "#d53e4f","#FF4E50", "#FC913A", "#fdae61",
           "#F9D423", "#fee08b" , "#abdda4" , "#a6d96a" , "#66c2a5" ,
           "#66bd63","#3B8183", "#3288bd", "#74add1",  "#abd9e9")
```


```{r}
### DAVIDE

file_list <- list.files("data")
data_list <- glue("data/{file_list}")

lga_data <- read_csv(data_list[1])
postcode_data <- read_csv(data_list[2])
age_data <- read_csv(data_list[4]) %>% filter(diagnosis_date >= "2021-05-01" & diagnosis_date <= "2021-09-30")
daily_lga_data <- read_csv(data_list[5]) %>% filter(diagnosis_date >= "2021-05-01" & diagnosis_date <= "2021-09-30")
source_data <- read_csv(data_list[7]) %>% filter(diagnosis_date >= "2021-05-01" & diagnosis_date <= "2021-09-30")

postcode_map_data <- read_sf("data/maps/POSTCODE_POLYGON.shp")
lga_map_data <- read_sf("data/maps/AD_LGA_AREA_POLYGON.shp") %>% filter(STATE == "VIC")

full_dataset <- read_csv(data_list[5]) 

all_lgas <- unique(full_dataset$Localgovernmentarea) %>% as.data.frame()
present_lgas <- unique(daily_lga_data$Localgovernmentarea) %>% as.data.frame()
missing_lgas <- all_lgas %>% filter(!. %in% present_lgas$.)%>% 
  mutate(Localgovernmentarea = gsub("\\s*\\([^\\]+\\)","",as.character(.)),
         Localgovernmentarea = toupper(Localgovernmentarea)) %>%
  select(-.) %>% 
  rbind("BULOKE", "COLAC OTWAY", "FRENCH ISLAND")

age_table <- table(age_data) %>% 
  as.data.frame() %>% group_by(diagnosis_date) %>% summarise(cases = sum(Freq)) %>%
  mutate(diagnosis_date = as.Date(diagnosis_date))
daily_lga_table <- table(daily_lga_data) %>% 
  as.data.frame() %>% 
  mutate(diagnosis_date = as.Date(diagnosis_date),
         Localgovernmentarea = gsub("\\s*\\([^\\]+\\)","",as.character(Localgovernmentarea)),
         Localgovernmentarea = toupper(Localgovernmentarea))  

source_table <- table(source_data) %>% 
  as.data.frame() %>% 
  mutate(diagnosis_date = as.Date(diagnosis_date))
```

```{r}
### IBRAHIM

file_list <- list.files("data")
data_list <- glue("data/{file_list}")

age_data <- read_csv(data_list[4]) %>%
  arrange(diagnosis_date) %>% 
  filter(diagnosis_date >= "2021-05-01" & diagnosis_date <= "2021-09-30") %>% 
  mutate(month = month(diagnosis_date, label = TRUE, abbr = TRUE),
         week = week(diagnosis_date),
         wday = wday(diagnosis_date, label = TRUE, abbr = TRUE),
         .after = diagnosis_date,
         agegroup = str_remove(agegroup, "_+$")) %>% 
  group_by(diagnosis_date) %>% 
  mutate(daily_count = n()) %>% 
  ungroup()
  

daily_lga_source_data <- read_csv(data_list[6]) %>%
  arrange(diagnosis_date) %>% 
  filter(diagnosis_date >= "2021-05-01" & diagnosis_date <= "2021-09-30") %>% 
  mutate(month = month(diagnosis_date, label = TRUE, abbr = TRUE),
         week = week(diagnosis_date),
         wday = wday(diagnosis_date, label = TRUE, abbr = TRUE),
         .after = diagnosis_date,
         Localgovernmentarea = str_remove(Localgovernmentarea, " \\(.*")) %>% 
  group_by(diagnosis_date) %>% 
  mutate(daily_count = n() ) %>% 
  ungroup()

week_age <- age_data %>% 
  filter(month == "Sep") %>% 
  mutate(week = case_when(week == 35 ~ 1,
                          week == 36 ~ 2,
                          week == 37 ~ 3,
                          week == 38 ~ 4,
                          week == 39 ~ 5)) %>% 
  count(week, agegroup)
```


class: center, middle

# How are the COVID-19 cases spreading in the Victorian community during 1st May to 30th September 2021?

???



---
class: center, middle

# Victoria COVID Maps

```{r, fig.align="center", out.height= "50%", out.width="75%"}
missing_lga_zero <- missing_lgas %>% mutate(total_cases = 0)

total_lga_cases <- daily_lga_table %>% 
  group_by(Localgovernmentarea) %>% 
  summarise(total_cases = sum(Freq)) %>%
  rbind(missing_lga_zero)

total_lga_map <-  left_join(lga_map_data, total_lga_cases, by = c("NAME" = "Localgovernmentarea"))

p3 <- ggplot(total_lga_map, aes(fill = total_cases))+
  geom_sf() + 
  theme_map() +
  labs(title = "Total New Cases",
       subtitle = "May 1 to September 30",
       fill = "Total Cases")+
  scale_fill_gradient(label = scales::label_comma()) +
  theme(legend.position = "right")

p3
```
---
class: center, middle

# Cases by Month 
---
class: center, middle

# May
```{r, fig.align="center", out.height= "50%", out.width="75%"}
months <- c("May", "Jun", "Jul", "Aug", "Sep")
missing_lga_month <- unique(missing_lgas$Localgovernmentarea) %>% expand.grid(months) %>% 
  mutate(total_cases = 0,
         Localgovernmentarea = as.character(Var1),
         month =as.character(Var2)) %>%
  select(Localgovernmentarea, month, total_cases)


monthly_lga_cases <- daily_lga_table %>% mutate(month = month(diagnosis_date, label = TRUE, abbr = TRUE)) %>%
  group_by(Localgovernmentarea, month) %>% summarise(total_cases = sum(Freq)) %>%
  rbind(missing_lga_month)

monthly_lga_map <-  left_join(lga_map_data, monthly_lga_cases, by = c("NAME" = "Localgovernmentarea"))

may_lga_map <- monthly_lga_map %>% filter(month == "May")

p4_may <- ggplot(may_lga_map, aes(fill = total_cases))+
  geom_sf() + 
  theme_map() +
  labs(title = "Total New Cases",
       subtitle = "May",
       fill = "Total Cases")+
  scale_fill_gradient(label = scales::label_comma()) +
  theme(legend.position = "right")

p4_may
```
---
class: center, middle

# June

```{r, fig.align="center", out.height= "50%", out.width="75%"}
Jun_lga_map <- monthly_lga_map %>% filter(month == "Jun")

p4_Jun <- ggplot(Jun_lga_map, aes(fill = total_cases))+
  geom_sf() +
  theme_map() +
  labs(title = "Total New Cases",
       subtitle = "June",
       fill = "Total Cases")+
  scale_fill_gradient(label = scales::label_comma()) +
  theme(legend.position = "right")

p4_Jun
```

---
class: center, middle

# July

```{r, fig.align="center", out.height= "50%", out.width="75%"}
Jul_lga_map <- monthly_lga_map %>% filter(month == "Jul")

p4_Jul <- ggplot(Jul_lga_map, aes(fill = total_cases))+
  geom_sf() +
  theme_map() +
  labs(title = "Total New Cases",
       subtitle = "July",
       fill = "Total Cases")+
  scale_fill_gradient(label = scales::label_comma()) +
  theme(legend.position = "right")

p4_Jul
```


---
class: center, middle

# August

```{r, fig.align="center", out.height= "50%", out.width="75%"}
Aug_lga_map <- monthly_lga_map %>% filter(month == "Aug")

p4_Aug <- ggplot(Aug_lga_map, aes(fill = total_cases))+
  geom_sf() +
  theme_map() +
  labs(title = "Total New Cases",
       subtitle = "August",
       fill = "Total Cases")+
  scale_fill_gradient(label = scales::label_comma()) +
  theme(legend.position = "right")

p4_Aug
```

---
class: center, middle

# September

```{r, fig.align="center", out.height= "50%", out.width="75%"}
Sep_lga_map <- monthly_lga_map %>% filter(month == "Sep")

p4_Sep <- ggplot(Sep_lga_map, aes(fill = total_cases))+
  geom_sf() +
  theme_map() +
  labs(title = "Total New Cases",
       subtitle = "September",
       fill = "Total Cases")+
  scale_fill_gradient(label = scales::label_comma()) +
  theme(legend.position = "right")

p4_Sep
```
---
class: center, middle

# Other findings!
---
class: center, middle, full-slide-fig

# Age
```{r}
age_data %>% ggplot(aes(agegroup)) +
  geom_bar(fill = "seagreen4") +
  theme_classic() +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(x = "Age Group",
       y = "Cases",
       title = "Total Cases by Age Group")
```

---
class: center, middle, full-slide-fig

```{r, fig.width=12}
age_data %>% ggplot(aes(agegroup)) +
  geom_bar(fill = "seagreen4") + 
  facet_wrap(vars(month)) +
  theme_classic() +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(x = "Age Group",
       y = "Cases",
       title = "Total Cases by Age Group per Month") 
```

---
class: center, middle

```{r}
age_data %>% count(month, name = "Cases") %>% 
  rename("Month" = month) %>% 
  kable(format.args = list(big.mark = ",")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---

class: center, middle

```{r, fig.width=12}
week_age %>% 
  ggplot(aes(agegroup, n)) +
  geom_col(fill = "seagreen4") +
  facet_wrap(vars(week)) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme_classic() +
  labs(x = "Age Group",
       y = "Cases",
       title = "Total Cases by Age Group per Week in September")
```

---
class: center, middle

```{r, fig.width=10}
week_age %>% group_by(week) %>% 
  summarise(week_total = sum(n)) %>% 
  mutate(weekly_inc_perc = round(((week_total / lag(week_total, default = week_total[1])) - 1) * 100),
         tot_inc_perc = round(week_total / week_total[1] * 100)) %>% 
  ggplot(aes(week, weekly_inc_perc)) +
  geom_line(color = "seagreen4", size = 1) +
  geom_label(aes(label = glue("{scales::comma(week_total)} cases")), fill = "mediumaquamarine", show.legend = FALSE) +
  theme_classic() +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(x = "Week",
       y = "Percentage Increase (%)",
       title = "Cases Percentage Increase per Week in September")
```

---
class: center, middle

```{r}
age_data %>% filter(month == "Sep") %>%
  ggplot(aes(wday)) +
  geom_bar(fill = "seagreen4") +
  theme_classic() +
  scale_y_continuous(labels = scales::label_comma()) +
  labs(x = "Weekday",
       y = "Cases",
       title = "Total Cases by Weekday in September")
```

---
class: center, middle

```{r}
age_data %>% filter(month == "Sep") %>% 
  group_by(wday) %>% 
  summarise(uni = n_distinct(diagnosis_date)) %>% 
  kable(col.names = c("Weekday", "Count")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---
class: center, middle

# LGA

```{r, fig.width=9}
highest_lga <- daily_lga_source_data %>% count(Localgovernmentarea, sort = TRUE) %>% 
  slice_head(n = 10) %>% pull(Localgovernmentarea)

daily_lga_source_data %>%
  filter(Localgovernmentarea %in% highest_lga) %>% 
  count(Localgovernmentarea, acquired) %>% 
  ggplot(aes(n, reorder(Localgovernmentarea, n, sum), fill = acquired)) +
  geom_col() +
  theme_classic() +
  theme(panel.grid.major.x = element_line(color = "grey", linetype = "dashed"),
        legend.position = c(.65, .2)) +
  scale_x_continuous(labels = scales::label_comma(), 
                     breaks = c(0, 1000, 2000, 5000)) +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "Cases",
       y = "Local Government Area",
       title = "Top Ten LGAs by Total Cases",
       fill = "Source")
```

---
class: center, middle

```{r, fig.width=9}
daily_lga_source_data %>% 
  filter(month == "Sep") %>% 
  count(diagnosis_date, acquired) %>% 
  ggplot(aes(diagnosis_date, n, color = acquired)) +
  geom_line(size = 1) +
  theme_classic() +
  theme(legend.position = "bottom") +
  scale_color_viridis(discrete = TRUE) +
  labs(x = "Date",
       y = "Cases",
       title = "Cases Trend per Source in September",
       color = "Source")
```

---
class: center, middle

```{r, fig.width=12}
daily_lga_source_data %>% 
  filter(month == "Sep",
         Localgovernmentarea %in% highest_lga) %>%
  count(diagnosis_date, Localgovernmentarea, acquired) %>% 
  ggplot(aes(diagnosis_date, n, color = acquired)) +
  geom_line() +
  facet_wrap(vars(Localgovernmentarea), scales = "free_y") +
  geom_line(size = 1) +
  theme_classic() +
  theme(legend.position = "bottom") +
  scale_color_viridis(discrete = TRUE) +
  labs(x = "Date",
       y = "Cases",
       title = "Cases Trend per Source per LGA in September",
       color = "Source")
```

---
class: center, middle

```{r, fig.width=12}
daily_lga_source_data %>% 
  filter(month == "Sep") %>% 
  mutate(week = case_when(week == 35 ~ 1,
                          week == 36 ~ 2,
                          week == 37 ~ 3,
                          week == 38 ~ 4,
                          week == 39 ~ 5)) %>% 
  count(diagnosis_date, week, acquired) %>% 
  ggplot(aes(diagnosis_date, n, color = acquired)) +
  geom_line() +
  facet_wrap(vars(week), scales = "free") +
  geom_line(size = 1) +
  theme_classic() +
  theme(legend.position = "bottom") +
  scale_color_viridis(discrete = TRUE) +
  labs(x = "Date",
       y = "Cases",
       title = "Cases Trend per Source per Week in September",
       color = "Source")
```

---
class: center, middle

# Thanks For Listening!

